---
title: "Statistical Inference Course Project - Part 2"
author: "Diego Dell'Era"
output: pdf_document
---

Tooth Growth
============



State your conclusions and the assumptions needed for your conclusions.

# Main goal

```
Analyze the ToothGrowth dataset. Perform a basic inferential data analysis.
```

# Exploratory analysis

### Load the dataset

```{r, results='markup'}
data(ToothGrowth)
summary(ToothGrowth)
```

The dataset analyzes the effect of vitamin C on tooth growth in guinea pigs. 

It contains `r dim(ToothGrowth)[1]` observations.

```
[,1]  len   numeric  Tooth length                <- this variable depends on the other two variables 
[,2]  supp  factor   Supplement type (VC or OJ). <- independent variable
[,3]  dose  numeric  Dose in milligrams.         <- independent variable
```

### Plot the dataset, isolating the independent variables

First, tooth length as a function of the supplement type. Then, tooth length as a function of the dose.

```{r, results='markup', fig.width=10}
par(mfrow=c(1,2))
boxplot(len ~ supp,
	    data = ToothGrowth,
	    xlab = "supplement type",
	    ylab = "tooth length",
	    main = "Tooth growth | supplement type")
boxplot(len ~ dose,
	    data = ToothGrowth,
	    xlab = "dose (in ml.)",
	    ylab = "tooth length",
	    main = "Tooth growth | dose")
```

### Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.

```{r, results='markup'}
t.test(len ~ dose, data=ToothGrowth)
```
